Sketch Recognition using Particle Swarm Algorithms

### 2009-02-11 trunk/smallpaper/SmallPaper.tex
before dr. magdah changes..

\documentclass[a4paper,10pt]{IEEEconf}
\usepackage[latin1]{inputenc}
\usepackage[american]{babel}
%\usepackage[T1]{fontenc}
\usepackage[dvips,pdftex]{graphicx}
%\usepackage{multicol}
\usepackage{subfigure}
\usepackage{named}
%\usepackage{fancybox}
%\usepackage{natbib}
\bibliographystyle{named}
\usepackage[dvips,pdftex]{hyperref}
%opening
\title{Sketch Recognition using Particle Swarm Algorithms}
%\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
\author{Maha El Meseery,  Mahmoud Fakhr El Din,  Samia Mashali,
\\  Magda Fayek and Nevin Darwish}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated. This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. This paper uses Support Vector Machines(SVM) classifier which further improves the final recognition accuracy.  %u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
 %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general. Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems. In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation. 
\end{abstract}
\section{Introduction}
%A picture speak a thousand words is also true for sketches. 
Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. %Engineers use paper and pencil design in early stages of design. 
Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}. %Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system. 
Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding. The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation. In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments. The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system. Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.

%the basic symbols are identified in consistent with the whole sketch. \\
%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated. Finally, the basic symbols are interpreted in consistent with the whole sketch. \\
This paper introduces a sketch recognition system. The presented system uses particle swarm algorithm (PSO) to optimally segment users strokes. Users can draw symbols using any number of strokes in any order they like. The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer. The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy. 
%The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.  Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system. The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.  The experiments and results achieved are explained in Section \ref{sec:Experiments}. Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work. 
%The next section will relate previous work in the field. The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm. Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details. Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
\section{Literature Review}
\label{Sec:RelatedWork}
%%%What is sketch recognitions. Why we need sketch recognition systems. 
The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition.
%To segment strokes
%In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.  %\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}. 
A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set. Their system has a draw back as it only recognizes simple specific shapes with a set of low level recognizers. \citeauthor{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area. Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10}. \citeauthor{Paleosketch08} \cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.

@@

Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. \citeauthor{CruveDivisionSwarm} uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve. \citeauthor{PolygonApproximationPSO} \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts \citeauthor{PolygonApproximationPSO} method but improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.% The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms. % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}

@@

The following sketch recognitions systems is divided into three main system blocks are 1) preprocessing, 2) segmentation and 3) Recognition. The following sections provides a detail description of each step.

 %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
%The preprocessing phase is responsible of capturing the input data and removing the noise from it. The system then proceeds to estimate a set of possible dominant points to help in the segmentation process. % The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data. The system then proceed to estimate a set of possible dominant points to help in segmentation process.  \\
  %then passes through the segmentation process which will try to

 %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
%The segmentation is divided into two steps ellipse fitting and curve segmentation. %In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms generate two segmentations. The system then chooses the segmentation that has the minimum segmentation error.%generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting

 %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.  The segmentation is then added to the set of un-recognized segments in the system. \\% This part is repeated for each stroke. 
 %The clustering algorithm starts to group segments together after the segmentation step. The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it. A set of hybrid features of statistical, geometric and spatial features is used to generate the symbol feature vector. The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
  %  Or to determine the symbol of the given segments from the set of preciously trained symbols. 
 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\ 
%\begin{figure}[]
%%\begin{minipage}[b]{0.5\linewidth}
%	\centering
%		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
%	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification. 
%	
%	\label{fig:Blockdiagram}
%%	\end{minipage}
%\end{figure}
\subsection{Preprocessing}
\label{Prepross}
%Some data is extracted from the points in the stroke.
 %It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines. Therefore, the 
 Speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}.  The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.  
 %speed is calculated as

@@

 %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
 %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
 %It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points.

@@

After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures. The system adopted the \textsl{average thresholding} process presented by \cite{earlyprocess}. This process is used to curvature, time difference and speed curves to generate a set of possible dominate points$P_{pd}$. These points are saved into a single array to be used as input to guide the next segmentation phase. %All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. They are then used as input to guide the next segmentation phase.%Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted the \textsl{average threshold} process presented by \cite{earlyprocess}. This process is used to curvature, time difference and speed curves. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. They are then used as input to guide the next segmentation phase.  %, where the mean

@@

After computing the primary data the system tries to segment the stroke into a set of primitives. The segmentation algorithm first attempts to fit the stroke points into a curve or an ellipse uisng a minimum square error fitting algorithm. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves (see Fig. \ref{fig:Blockdiagram}). The next section describes the two particle swarm algorithms used to divide the stroke.

@@

%\subsubsection{Ellipse Detection} 

@@

%The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation. %After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.

@@

Two DPSO algorithms are used to generate segmentations for each stroke the user draws. The segmentation with the minimum error value is chosen as the stroke segmentation. The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions.  
\paragraph{Problem definition}
The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set 
$A  = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem with binary array with the same length $N$ as the input stroke.  

@@

Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1. When $p_{ij}=1$ means that point $j$ is a dominant point. 
\paragraph{The fitness functions}

@@

The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 

@@

\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the value of solution. Otherwise the system favors the lower number of vertices.

@@

\section{Experiments}\label{sec:Experiments}
A dataset collected by \citeauthor{HeloiseBeautification} is used to test the presented system\cite{HeloiseBeautification}. The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. We divide the dataset into four sets each set contains four different users. Then we divide the dataset equally into training set and test set. To generate trusted results, four different splits for the training and test data are generated from the dataset. The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. In the following section, the recognition accuracy means the average recognition accuracy in the four data splits. %The result presented in this section are the average result of the four splits. %then result are stored The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. %Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged. %We divide the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\

@@

\begin{figure}[]\centering
\fbox{ \parbox{3.5cm}{% 
		\includegraphics[scale=0.4]{images/symbolSet.PNG}	}}
	\caption{The Symbol Set} Ths figure shows a set of the symbols used in the dataset. 
	\label{fig:symbolSet}

@@

		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/result1.jpg} 	\label{fig:test1}}
		\hfill
  \subfigure [Symbols Comparison ] {\includegraphics[width=10.5cm]{images/result2.jpg} \label{fig:test2}}
	\caption{Experiments results:}  a)The results of comparing the algorithms. The graphs displays the recognition rate of symbols. b) The graph shows the recognition rate of each symbol using different algorithms.
\end{figure*}
%In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). %We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
  Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms. The swarm algorithms were tested with and without the ellipse fitting module. The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.
The second experiment we implemented is testing the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. Figure \ref{fig:test2} also indicates that algorithm \textsl{AlgS1} alone achieve better performance than algorithm   \textsl{AlgS2} in the symbols that consist only of lines. This is natural as algorithm  \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. %Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. 

@@

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}The next step in 
A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. The clustering must be performed without the user explicit involvement. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol. 
%\bibliography{Mybibliography}
\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
\end{document}

>>>>

\documentclass{article}%{IEEEconf}[a4paper,10pt]
\usepackage{spconf}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage[american]{babel}
%\usepackage[T1]{fontenc}
\usepackage[dvips,pdftex]{graphicx}
%\usepackage{multicol}
\usepackage{subfigure}
\usepackage{named}
%\usepackage{fancybox}
%\usepackage{natbib}

%\bibliographystyle{named}
\usepackage[dvips,pdftex]{hyperref}
%opening
\title{Sketch Recognition using Particle Swarm Algorithms}
%\title{A Study on Using Particle Swarm Algorithms in Symbol Recognition }
%\author{Maha El Meseery ;  Mahmoud Fakhr El Din ;  Samia Mashali ;  Magda Fayek and Nevin Darwish }
%\name{Maha El Meseery ;  Mahmoud Fakhr El Din ;  Samia Mashali ;  Magda Fayek and Nevin Darwish }

\twoauthors
  {Maha El Meseery, Mahmoud Fakhr El Din, Samia Mashali}
  %{}
	{Signals Processing Group \\ Computers and Systems Department \\
  Electronic Research Institute\\Cairo, Egypt\\
	melmeseery@eri.sci.eg, samia@eri.sci.eg}
  {Magda Fayek, Nevin Darwish}%\sthanks{The fourth author performed the work while at ...}
  %{}
	{Department of Computer Engineering\\
	Faculty of Engineering \\Cairo University\\Cairo, Egypt\\
	 magda@gmail.com,ndarwish@ieee.org}

%\address{ Electronic research inistiute}
\begin{document}
\maketitle
\begin{abstract}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated. This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. This paper uses Support Vector Machines(SVM) classifier which further improves the final recognition accuracy.  %u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
 %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general. Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems. In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation. 
\end{abstract}
\begin{keywords}
Handwriting recognition, pattern recognition ,Document image processing
\end{keywords}
\section{Introduction}
%A picture speak a thousand words is also true for sketches. 
The field of symbol recognition has gained interest in the last few years. A wide variety of techniques were used either on segmentation process or symbol recognition. Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. %Engineers use paper and pencil design in early stages of design. 
Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems. %Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system. 
Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition. The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation. In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments. The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system. Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols.

%the basic symbols are identified in consistent with the whole sketch. \\
%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated. Finally, the basic symbols are interpreted in consistent with the whole sketch. \\
This paper introduces a sketch recognition system. The presented system uses particle swarm algorithm (PSO) to optimally segment users strokes. Users can draw symbols using any number of strokes in any order they like. The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer. The experiments in section \ref{sec:Experiments} shows that the proposed system can improve the final symbol recognition accuracy. 
%The remaining of this papers is organized as following; section \ref{Sec:RelatedWork} provides a brief literature review.  Section \ref{PSO} explains the general particle swarm algorithm which is used in the presented system. The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}.  The experiments and results achieved are explained in Section \ref{sec:Experiments}. Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work. 
%The next section will relate previous work in the field. The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm. Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details. Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\
%\section{Literature Review}
%\label{Sec:RelatedWork}
%%%What is sketch recognitions. Why we need sketch recognition systems.

%To segment strokes
%In order to segment strokes, \cite{meanshift10} introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on this area.  %\citeauthor{HeloiseBeautification} used a dynamic programming algorithm to solve the segmentation problem \cite{HeloiseBeautification}. 
A hybrid algorithm was introduced in \cite{earlyprocess} where they generated different sets of segmentation based on both curvature and speed dominant points, this is followed by choosing a segmentation with the least error from generated hybrid set. Their system has a draw back as it only recognizes specific simple geometric shapes with a set of low level recognizers. Yu introduced a feature area for each primitive and then computed the segmentation error for different types of primitives based on the computed feature area. Their system achieved good accuracy in simple shapes (square, ellipse,...) but did not perform well in more complex shapes\cite{meanshift10}. Paulson and Hammond\cite{Paleosketch08} introduced a set of low level recognizers that were reported to achieve 98\%  accuracy but their system similar to all low level recognizers only identify a small set of simple shapes.

@@

Genetic algorithm was used by \cite{CruveDivisionSwarm} to optimally divide digital curves into lines and curves. Chen et al. uses digital curves scanned from paper as input to the system and did not take advantage of the curvature or local geometric properties of the digital curve. Yin \cite{PolygonApproximationPSO} used PSO to convert digital curves into polygons, our system adopts Yin's method and improves it by adding curvature and other local information while segmenting strokes to achieve better segmentations.% The next section presents the general particle swarm algorithm which is used in our proposed segmentation algorithms. % are presented in section \ref{subsubsec:Discreteparticleswarmalgorithm}

@@

The following sketch recognitions systems is divided into three main system steps are 1) preprocessing, 2) segmentation and 3) Recognition. The following sections provides a detail description of each step.
 %Due to sloppiness of users and hardware glitches the system preprocess the data before further processing. 
%The preprocessing phase is responsible of capturing the input data and removing the noise from it. The system then proceeds to estimate a set of possible dominant points to help in the segmentation process. % The time differences, direction, speed and curvature of all points of the stroke is computed along with some other geometrical data. The system then proceed to estimate a set of possible dominant points to help in segmentation process.  \\
  %then passes through the segmentation process which will try to 
 %The goal of the segmentation stage is to divide stroke into segments of either curves or lines.
%The segmentation is divided into two steps ellipse fitting and curve segmentation. %In the first step, the ellipse fitting process tries to fit the stroke into an ellipse. If the system fails, it passes the stroke to the second step; curve segmentation which consists of two PSO segmentation algorithms. The two algorithms generate two segmentations. The system then chooses the segmentation that has the minimum segmentation error.%generate the segmentations,  the minimum error will be the chosen segmentation. % is the ellipse fitting
 %the stroke first If the ellipse detection fails the stroke is passed to the segmentation algorithms which will pass it to the two PSO algorithms described below the segmentation with the minimum error will be the chosen segmentation.  The segmentation is then added to the set of un-recognized segments in the system. \\% This part is repeated for each stroke. 
 %The clustering algorithm starts to group segments together after the segmentation step. The system allows the user draws the symbol by using any number of strokes; a set of unrecognized segments is passed to the clustering algorithm to generate a symbol and compute a feature vector for it. A set of hybrid features of statistical, geometric and spatial features is used to generate the symbol feature vector. The final classification step, SVM classifier uses the feature vector computed to classify the segments into one of the previously trained classes.
  %  Or to determine the symbol of the given segments from the set of preciously trained symbols. 
 %with a symbol from the training set.% The system compute composite set of features some are statistical other are spatial features based on the type primitives. Finally, the strokes are classified into the corresponding classes.% the classifier identifies the strokes into a symbol from the set of known symbols.\\ 
%%\begin{minipage}[b]{0.5\linewidth}
%		\includegraphics[scale=0.75]{images/block2.pdf}%Blockdiagram.pdf}
%	\caption{The system Block Diagram.} The figure shows the main system blocks. Preprocessing, segmentation, clustering, features extraction and classification. 
%	\label{fig:Blockdiagram}
%%	\end{minipage}
\subsection{Preprocessing}
\label{Prepross}
%Some data is extracted from the points in the stroke.
 %It is noted that as the user draws a shape, the pen slows down near corners and picks up speed when drawing a straight lines. Therefore, the 
 Speed and time difference data were widely used in sketch understanding systems \cite{earlyprocess}. The direction and curvature data are used to determine the points with high angular changes along the path of points \cite{meanshift10}.  
 %speed is calculated as 

@@

 %system compute speed, distance data for all points in the stroke. Curvature is computed using estimation used in [] where direction computed is angle between two lines. Curvature is 
 %$\Delta d/  \Delta S $ where d is the difference direction of point and s is difference in distance between points. 
 %It should be noted that, all these calculations are performed in real time while the user draws the strokes. The complexity of computation is $O(n)$ where $n$ is number of points.

@@

After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures. Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated and used as threshold \textit{th}. The threshold \textit{th} is used to separate the curve into regions  $Region_i$; each region $Region_i$ is defined as the part between two intersection points of the threshold line and the curve. Regions higher than the threshold \textit{th} are processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. The system repeats this process to curvature, time difference and speed curves. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. %They are then used as input to guide the next segmentation phase. 
%After the system computes all the speed, time difference and curvature data it proceeds to detect the points with low velocity and high curvatures. The system adopted the \textsl{average thresholding} process presented by \cite{earlyprocess}. This process is used to curvature, time difference and speed curves to generate a set of possible dominate points$P_{pd}$. These points are saved into a single array to be used as input to guide the next segmentation phase. %All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. They are then used as input to guide the next segmentation phase.%Using simple differentiations to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted the \textsl{average threshold} process presented by \cite{earlyprocess}. This process is used to curvature, time difference and speed curves. All the points labeled as possible dominant points $P_{pd}$ are saved into a single array. They are then used as input to guide the next segmentation phase.  %, where the mean

@@

After computing the primary data the system tries to segment the stroke into a set of primitives. The segmentation algorithm first attempts to fit the stroke points into a curve or an ellipse uisng a minimum square error fitting algorithm. If the stroke proved to be an ellipse arc then the segmentation process ends and the system proceeds to the next step. Otherwise, the stroke is passed into two particle swarm algorithms that divides the stroke to either lines or lines and curves. The algorithms take the stroke points along with the possible dominant points $P_{pd}$ computed then produce a set of dominant points which are connected with either lines or curves. The next section describes the ellipse detection algorithm and the two particle swarm algorithms used to divide the stroke.\\ %The next section describes the two particle swarm algorithms used to divide the stroke. %(see Fig. \ref{fig:Blockdiagram})

@@

 \begin{figure}
		\includegraphics[scale=0.7]{images/segblock.jpg}
	\caption{Segmentation Block}
	\label{fig:segblock}
\subsubsection{Ellipse Detection} 

@@

The process starts by computing the center of the stroke bounding box. The bounding box center point is used as the first estimation of the ellipse center. The axes of the ellipse are estimated as $width/2$ and $height/2$ of the stroke bounding box. The least square fitting algorithm is used to minimize the fitting error of the ellipse equation. %After few loops, the error and a confidence values  are computed to check if the stroke can be labeled as ellipse or not.

@@

Two DPSO algorithms are used to generate segmentations for each stroke the user draws. The segmentation with the minimum error value is chosen as the stroke segmentation. The problem definition is the same in both algorithms but they differ in the method they use to compute fitness and error functions. \\
%\paragraph{Problem definition}
\textsl{Problem definition:} The input stroke with $N$ points can be represented by set $S = \left\{ {x_1 ,x_2  \ldots x_N }\right\}$ where $x_i$ is the location of the point $i$. The swarm algorithms consist of $M$ agents which are represented by the set  $A = \left\{ {P_i \left| {i = 1,2 \cdots M} \right.} \right\}$ where $P_i$ is a single solution particle from the solution space. Each particle decodes the problem with binary array with the same length $N$ as the input stroke.  

@@

Therefore, the system represents each particle $P_i$ by $P_i = \left\{ {p_{ij} \left| {j = 1,2 \cdots N} \right.} \right\}$ where $p_{ij}$ has only two values 0 or 1. When $p_{ij}=1$ means that point $j$ is a dominant point. \\
\begin{figure}
			\subfigure[Particle Encoding]{	\includegraphics[scale=0.5]{images/pso1.jpg}}
					\subfigure[Posible domninate point]{	\includegraphics[scale=0.5]{images/ppd.jpg}}
	\caption{}
	\label{fig:pso1}
%\paragraph{The fitness functions}

@@

\textsl{Fitness function:} The fitness function and error calculation are different in each of the two \textit{DPSO} algorithms. For the first algorithm \textsl{AlgS1}, 

@@

\end{equation} where $N$ is the number of points in the stroke, $D$ is the number of points in the solution that was previously labeled as a possible dominant point ($P_{pd}$), $E$ is the computed error and $\varepsilon$ is the error threshold. It should be noticed that when the error is larger than the threshold $E$ the fitness is given a -ve value to lower the fitness value of the solution. Otherwise the system favors the lower number of vertices.

@@

\section{Experiments}
\label{sec:Experiments}
The dataset collected by Hse and Newton is used to test the presented system\cite{HeloiseBeautification}. The data are drawn by 16 users each of them was asked to draw the 13 shapes from 30 to 50 times. Figure \ref{fig:symbolSet} shows a set of the shapes used in the data set. We divide the dataset into four sets each set contains four different users. Then we divide the dataset equally into training set and test set. To generate trusted results, four different splits for the training and test data are generated from the dataset. The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. In the following section, the recognition accuracy means the average recognition accuracy in the four data splits. %The result presented in this section are the average result of the four splits. %then result are stored The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. %Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged. %We divide the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\

@@

%\begin{figure}[]\centering
%	\caption{The Symbol Set} 
%	\label{fig:symbolSet}

@@

	\begin{minipage}[b]{4cm}
	\subfigure [The Symbol Set] {\fbox{ \parbox{2cm}{% 
		\includegraphics[scale=0.2]{images/symbolSet.PNG}	}}}

		\subfigure[Algorithm comparison]{\includegraphics[width=5cm]{images/result1.jpg} 	\label{fig:test1}}
		\end{minipage}
		\hfill
  \subfigure [Symbols Comparison ] {\includegraphics[width=10.5cm]{images/result2.jpg} \label{fig:test2}}
	\caption{Experiments results:} a) The figure shows a set of the symbols used in the dataset.   b)The results of comparing the algorithms. The graphs displays the recognition rate of symbols. c) The graph shows the recognition rate of each symbol using different algorithms. 
\end{figure*}
%In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets .

We performed two experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). %We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to our swarm algorithms.
  Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The results shows that both \textit{DPSO} algorithms achieve better result than other algorithms. The swarm algorithms were tested with and without the ellipse fitting module. The ellipse detection module appears to be superior to results with only the \textit{DPSO} algorithms.

The second experiment we implemented was testing the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows the accuracy of each symbol, it is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. Figure \ref{fig:test2} also indicates that algorithm \textsl{AlgS1} alone achieve better performance than algorithm   \textsl{AlgS2} in the symbols that consist only of lines. This is natural as algorithm  \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. %Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of lines and curves, if the stroke consists of curve only  the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. 

@@

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}The next step in 
A possible extension of this research is to complete the clustering algorithm for fully automated sketch recognition. The clustering must be performed without the user explicit involvement. Other area of enhancements is the features extraction methods. Introducing more spatial and geometrical features is believed to improve classifications.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

%\bibliography{Mybibliography}
\bibliographystyle{IEEEbib}
\bibliography{../../neededfiles/Bibliographies/Mybibliography2}
\end{document}
